---
title: "Exploratory Data Analysis in R"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: show
    number_sections: true
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r init , include=FALSE}
# Loading necessary libraries

library(ezids)
```

```{r setup, include=FALSE}
# some of common options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
knitr::opts_chunk$set(warning = F, results = "markup", message = F)
# knitr::opts_chunk$set(warning = F, results = "hide", message = F)
options(scientific=T, digits = 3) 
```

#### Importing the necessary libraries
```{R}
library(dplyr)
library(tidyverse)  
library(lubridate)  
library(ggplot2)    
library(wordcloud)

```
```{r}
library(readxl)
df=read_xlsx("D:/Githubfordatamining/FA24-DATS6101-11/Project1/EDA_Global_Terrorism/Global_terrorism_data_cleaned_01.xlsx")

head(data)
str(data)

```
```{r}
data$success <- as.factor(data$success)
data$Group <- as.factor(data$Group)
data$AttackType <- as.factor(data$AttackType)
data$Target_type <- as.factor(data$Target_type)
data$Weapon_Type <- as.factor(data$Weapon_Type)

data$casualities <- as.integer(data$casualities)
data$Killed <- as.integer(data$Killed)
data$Wounded <- as.integer(data$Wounded)

str(data)
```


No of terrorist attacks over the time .
```{r}
# Load required library
library(ggplot2)

# Count the number of incidents by year
yearly_counts <- table(data$Year)

# Convert to a data frame
yearly_counts_df <- as.data.frame(yearly_counts)
colnames(yearly_counts_df) <- c("Year", "Total_Incidents")

# Convert Year to numeric for proper plotting
yearly_counts_df$Year <- as.numeric(as.character(yearly_counts_df$Year))

# Plotting the frequency of incidents over time as a line graph
ggplot(yearly_counts_df, aes(x = Year, y = Total_Incidents)) +
  geom_line(color = "steelblue", size = 1) +  # Line graph
  geom_point(color = "red", size = 2) +  # Optional: Add points at each year
  labs(title = "Frequency of Terrorist Activities Over Time",
       x = "Year", 
       y = "Number of Incidents") +
  scale_x_continuous(breaks = yearly_counts_df$Year) +  # Set x-axis to show all years
  theme_minimal() +  # Optional: A clean theme
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) 


```
Terrorist attacks by region 
```{r}
# Create a table of Year and Region counts from the sampled data
region_counts_sample <- table(data$Year,data$Region)
region_counts_sample_df <- as.data.frame(region_counts_sample)

# Rename columns for clarity
colnames(region_counts_sample_df) <- c("Year", "Region", "Freq")

region_counts_sample_df$Region <- as.factor(region_counts_sample_df$Region)

# Plotting with sampled data

library(ggplot2)
ggplot(region_counts_sample_df, aes(x = as.numeric(as.character(Year)), y = Freq, fill = Region)) +
  geom_area(position = 'stack') +
  labs(title = "Terrorist Activities by Region Over Time",
       x = "Year",
       y = "Number of Attacks") +
  theme_minimal()

```

Number of attacks happened with different types of terrorist attacks.
```{r}
# Count tactics used in attacks
tactics_count <- table(data$AttackType)

# Convert to a data frame and sort
tactics_count_df <- as.data.frame(tactics_count)
tactics_count_df <- tactics_count_df[order(-tactics_count_df$Freq),]

# Set margins to allow for long x-axis labels
par(mar = c(10, 5, 4, 2), mgp = c(3, 0.5, 0))  # Adjust bottom margin (first value) to make space

# Plotting the bar chart with adjusted label sizes and vertical orientation
barplot(tactics_count_df$Freq, 
        names.arg = tactics_count_df$Var1, 
        col = "darkorange", 
        las = 2,                # Rotate labels to be vertical
        cex.names = 0.6,        # Adjust the font size of the x-axis labels
        main = "Tactics Employed in Terrorist Attacks",
        ylab = "Number of Attacks")

```

Outcome of terrorist attacks over the world.
```{r}
# Assuming df is your dataframe and 'success' is the relevant column
# Group by 'success' and calculate the count of each outcome
succ <- as.data.frame(table(data$success))

# Rename the columns for clarity
colnames(succ) <- c("success", "count")

# Calculate the total number of attacks (sum of all counts)
sum_of_attacks <- sum(succ$count)

# Calculate the percentage for each outcome
succ$percentage <- (succ$count / sum_of_attacks) * 100

# Display the result
print(succ)

# Plot a bar chart showing the percentage of success
# Load necessary libraries
library(ggplot2)
ggplot(succ, aes(x = factor(success), y = percentage, fill = factor(success))) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  labs(title = "Outcome of Terrorist Attacks over the World",
       x = "Outcome",
       y = "Percentage") +
  theme_minimal()

```

```{r}
length(data$casualties)
length(data$Year)
```

```{r}
# Correlation between incidents and casualties
correlation <- cor(yearly_counts_df$Total_Incidents, 
                   tapply(data$casualities, data$Year, sum, na.rm = TRUE))

# Plotting the correlation
plot(yearly_counts_df$Total_Incidents, 
     tapply(data$casualities, data$Year, sum, na.rm = TRUE),
     main = "Correlation between Number of Incidents and Casualties",
     xlab = "Total Incidents", ylab = "Total Casualties")
abline(lm(tapply(data$casualities, data$Year, sum, na.rm = TRUE) ~ 
          yearly_counts_df$Total_Incidents), col = "red")
```
Top 10 countries affected.
```{r}
# Load required library
library(ggplot2)

# Count occurrences of each country
country_counts <- table(data$Country)  # Create a frequency table
top_countries <- sort(country_counts, decreasing = TRUE)  # Sort in descending order
top_countries <- top_countries[1:10]  # Get the top 15 countries

# Convert to a data frame for ggplot
top_countries_df <- data.frame(Country = names(top_countries), Count = as.vector(top_countries))

# Create a bar plot
ggplot(top_countries_df, aes(x = reorder(Country, Count), y = Count, fill = Count)) +
  geom_bar(stat = "identity", width = 0.5) +
  scale_fill_gradientn(colors = c("green", "orange", "red")) + # Gradient from green to orange to red
  labs(title = "Top Countries Affected", 
       x = "Countries", 
       y = "Count") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) + # Rotate x labels
  theme_minimal() + # Optional: A clean theme
  theme(legend.position = "none") 

```
Chi-square test to test the relationship between the Attack type and the sucess rate.

Null Hypothesis (H0): There is no association between AttackType and success.

Alternative Hypothesis (H1): There is an association between AttackType and success.

Since the p-value < 0.05 we reject the null hypothesis that there is a relationship between the attacktype and the sucess rate.
```{r}
contingency_table <- table(data$AttackType, data$success)

chi_squared_results <- chisq.test(contingency_table)
print(chi_squared_results)
```
Proportion Test
Null Hypothesis (H0): The proportion of successful attacks is the same in both countries.

Alternative Hypothesis (H1): The proportion of successful attacks is different in both countries.

p-value > 0.05 fail to reject the null hypothesis. The propotion of attack is same in both countries.
```{r}
# Create a subset for two countries
country_data <- data[data$Country %in% c("Afghanistan", "Iraq"), ]

# Create a contingency table for success
success_table <- table(country_data$Country, country_data$success)

# Perform the proportion test
prop_test_results <- prop.test(success_table)
print(prop_test_results)

```
T-test

Null Hypothesis (H0): There is no difference in the mean casualties between the two attack types.

Alternative Hypothesis (H1): There is a difference in the mean casualties between the two attack types.

Since p-value < 0.05 we reject the null hypothesis which concludes that there difference in the mean casualities for the two attack types.

```{r}

# Subset the data for two attack types
attack_data <- data[data$AttackType %in% c("Armed Assault", "Bombing/Explosion"), ]

# Perform a t-test
t_test_results <- t.test(casualities ~ AttackType, data = attack_data)
print(t_test_results)
```
Chi-sqaure test

Null Hypothesis (H₀): AttackType and Region are independent.

Alternative Hypothesis (H₁): There is an association between AttackType and Region.

Since p-value < 0.05 you reject the null hypothesis.



```{r}
attack_table <- table(data$AttackType,data$Region)

chisq.test(attack_table)
```
Cramér's V quantifies the strength of the association between AttackType and Region. A value of 0.139 indicates a weak association.

```{r}
library(vcd)

assoc_stats <- assocstats(attack_table)

assoc_stats
```

