---
title: "Data Cleaning and pre-processing"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: show
    number_sections: true
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r init , include=FALSE}
# Loading necessary libraries

library(ezids)
```

```{r setup, include=FALSE}
# some of common options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
knitr::opts_chunk$set(warning = F, results = "markup", message = F)
# knitr::opts_chunk$set(warning = F, results = "hide", message = F)
options(scientific=T, digits = 3) 

```
### Data Pre-processing

Reading the file and checking on the structure of the data.

```{r}
data <- read.xlsx("C:/Users/darsh/Desktop/Global_terrorism_raw_data.xlsx")

head(data)
str(data)

```
Checking on the column names in the dataset.

```{r}
column_names <- colnames(data)
print(column_names)
```
Renaming the columns for better readability.

```{r}
names_to_rename <- c("iyear" = "Year",
                      "imonth" = "Month",
                      "iday" = "Day",
                      "country_txt" = "Country",
                      "provstate" = "State",
                      "region_txt" = "Region",
                      "attacktype1_txt" = "AttackType",
                      "nkill" = "Killed",
                      "nwound" = "Wounded",
                      "gname" = "Group",
                      "targtype1_txt" = "Target_type",
                      "weaptype1_txt" = "Weapon_Type",
                      "motive" = "Motive")

# Loop through the names and rename the columns

for (old_name in names(names_to_rename)) {
  new_name <- names_to_rename[old_name]
  colnames(data)[colnames(data) == old_name] <- new_name
}
```

Selecting only the important columns from the dataset which is relevant for the analysis.
```{r}

data <- data[, c("Year", "Month", "Day", "Country", "State", "Region", 
              "AttackType", "Killed", "Wounded", "Group", "Target_type", 
              "Weapon_Type", "Motive","success")]

```

Checking if there is any NA values in the columns selected.

Checking the dim and structure of the data : 181691 rows and 15 columns.
```{r}
head(data)
colSums(is.na(data))

column_names <- colnames(data)
print(column_names)

dim(data)
str(data)
```

Including a calculative column.
```{r}
data$casualities <- data$Killed + data$Wounded

colnames(data)
```

Handling the NA values.

```{r}
# Calculate median for the 'Killed','Wounded'and 'casualities' columns

median_killed <- median(data$Killed, na.rm = TRUE)
median_wounded <- median(data$Wounded, na.rm = TRUE)
median_casualities <- median(data$casualities , na.rm = TRUE)

# Replacing the NA values with the median

data$Killed[is.na(data$Killed)] <- median_killed
data$Wounded[is.na(data$Wounded)] <- median_wounded
data$casualities[is.na(data$casualities)] <- median_casualities

data$Motive[is.na(data$Motive)] <- "Unknown"

sum(is.na(data$Killed))  
sum(is.na(data$Wounded)) 
sum(is.na(data$casualities))

colSums(is.na(data))
```
```{r}
str(data)
```
Changing the datatypes of the variables.

```{r}
data$success <- as.factor(data$success)
data$Group <- as.factor(data$Group)
data$AttackType <- as.factor(data$AttackType)
data$Target_type <- as.factor(data$Target_type)
data$Weapon_Type <- as.factor(data$Weapon_Type)

data$casualities <- as.integer(data$casualities)
data$Killed <- as.integer(data$Killed)
data$Wounded <- as.integer(data$Wounded)

str(data)
```
Saving the cleaned data set.

```{r}
library(writexl)
write_xlsx(data, "C:/Users/darsh/Documents/Global_terrorism_data_cleaned_01.xlsx")
```
```{r}
library(readxl)
data=read_xlsx("D:/Githubfordatamining/FA24-DATS6101-11/Project1/EDA_Global_Terrorism/Global_terrorism_data_cleaned_01.xlsx")
str(df)
```
```{r}
data$success <- as.factor(data$success)
data$Group <- as.factor(data$Group)
data$AttackType <- as.factor(data$AttackType)
data$Target_type <- as.factor(data$Target_type)
data$Weapon_Type <- as.factor(data$Weapon_Type)

data$casualities <- as.integer(data$casualities)
data$Killed <- as.integer(data$Killed)
data$Wounded <- as.integer(data$Wounded)

str(data)
```

```{r}
# Load required library
library(ggplot2)

# Count the number of incidents by year
yearly_counts <- table(data$Year)

# Convert to a data frame
yearly_counts_df <- as.data.frame(yearly_counts)
colnames(yearly_counts_df) <- c("Year", "Total_Incidents")

# Convert Year to numeric for proper plotting
yearly_counts_df$Year <- as.numeric(as.character(yearly_counts_df$Year))

# Plotting the frequency of incidents over time as a line graph
ggplot(yearly_counts_df, aes(x = Year, y = Total_Incidents)) +
  geom_line(color = "steelblue", size = 1) +  # Line graph
  geom_point(color = "red", size = 2) +  # Optional: Add points at each year
  labs(title = "Frequency of Terrorist Activities Over Time",
       x = "Year", 
       y = "Number of Incidents") +
  scale_x_continuous(breaks = yearly_counts_df$Year) +  # Set x-axis to show all years
  theme_minimal() +  # Optional: A clean theme
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) 


```
```{r}
# Create a table of Year and Region counts from the sampled data
region_counts_sample <- table(data$Year,data$Region)
region_counts_sample_df <- as.data.frame(region_counts_sample)

# Rename columns for clarity
colnames(region_counts_sample_df) <- c("Year", "Region", "Freq")

region_counts_sample_df$Region <- as.factor(region_counts_sample_df$Region)

# Plotting with sampled data

library(ggplot2)
ggplot(region_counts_sample_df, aes(x = as.numeric(as.character(Year)), y = Freq, fill = Region)) +
  geom_area(position = 'stack') +
  labs(title = "Terrorist Activities by Region Over Time",
       x = "Year",
       y = "Number of Attacks") +
  theme_minimal()

```
```{r}
# Count tactics used in attacks
tactics_count <- table(data$AttackType)

# Convert to a data frame and sort
tactics_count_df <- as.data.frame(tactics_count)
tactics_count_df <- tactics_count_df[order(-tactics_count_df$Freq),]

# Set margins to allow for long x-axis labels
par(mar = c(10, 5, 4, 2), mgp = c(3, 0.5, 0))  # Adjust bottom margin (first value) to make space

# Plotting the bar chart with adjusted label sizes and vertical orientation
barplot(tactics_count_df$Freq, 
        names.arg = tactics_count_df$Var1, 
        col = "darkorange", 
        las = 2,                # Rotate labels to be vertical
        cex.names = 0.6,        # Adjust the font size of the x-axis labels
        main = "Tactics Employed in Terrorist Attacks",
        ylab = "Number of Attacks")

```
```{r}
ggplot(df, aes(x = AttackType)) + 
  geom_bar(fill="#69b3a2",color="purple") + 
  theme_minimal()+
  labs(title = "Distribution of Attack types",
       x="Attack type",
       y="count")+
  scale_fill_brewer(palette = "Set3") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 


```
What are the most common types of attacks, and how do they vary by region or country?




```{r}
# Count the number of attacks by country and select the top 10
top_countries <- df %>%
  group_by(Country) %>%
  summarise(Count = n()) %>%
  top_n(10, Count)

# Plot the top 10 countries
ggplot(top_countries, aes(x = reorder(Country, -Count), y = Count)) + 
  geom_bar(stat = "identity", fill="lightblue", color="purple") + 
  labs(title = "Top 10 Countries with Most Terrorist Attacks",
       x="Country",
       y="Count of Attacks") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_brewer(palette = "Set3")

```
The above Visualization Which regions or countries have experienced the most attacks over the years, and how has that changed over time?
Here, the focus is on the top 10 countries with the highest number of attacks, providing a snapshot of where the most incidents have occurred.

```{r}

# Summarize attacks by month and type
attack_seasonality <- df %>%
  group_by(Month, AttackType) %>%
  summarise(attack_count = n())

# Heatmap to show seasonality
ggplot(attack_seasonality, aes(x = Month, y = AttackType, fill = attack_count)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "blue") +
  theme_minimal() +
  labs(title = "Seasonality of Terrorist Attacks", y = "Attack Type", x = "Month")



```
```{r}
colnames(df)
# Load necessary libraries
library(ggplot2)
library(dplyr)
library(reshape2)



# Grouping data by region and year to count the number of attacks
attacks_by_region_year <- df%>%
  group_by(Year, Region) %>%
  summarise(Attack_Count = n())

# Reshape data for heatmap
attacks_by_region_pivot <- dcast(attacks_by_region_year, Year ~ Region, value.var = "Attack_Count", fill = 0)

# Plotting heatmap
ggplot(melt(attacks_by_region_pivot, id.vars = 'Year'), aes(x = variable, y = Year, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "lightblue", high = "blue") +
  theme_minimal() +
  labs(title = "Terror Attacks by Region Over the Years", x = "Region", y = "Year") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))



```

```{r}

# Load necessary libraries
library(ggplot2)
library(dplyr)

# Summarize the number of attacks by country and year
country_trends <- df %>%
  group_by(Year, Country) %>%
  summarise(Attack_Count = n()) %>%
  arrange(desc(Attack_Count))

# Filter for the top 10 most affected countries over all years
top_countries <- country_trends %>%
  group_by(Country) %>%
  summarise(Total_Attacks = sum(Attack_Count)) %>%
  arrange(desc(Total_Attacks)) %>%
  top_n(10, wt = Total_Attacks)

# Filter the data for these top 10 countries
country_trends_filtered <- country_trends %>%
  filter(Country %in% top_countries$Country)

# Plot the trend over the years for the top 10 countries
ggplot(country_trends_filtered, aes(x = Year, y = Attack_Count, color = Country)) +
  geom_line(size = 1) +
  labs(title = "Terror Attacks Over the Years in the Most Affected Countries", 
       x = "Year", y = "Number of Attacks") +
  theme_minimal() +
  theme(legend.position = "bottom")



```
```{r}
# Summarize the total number of casualties by terrorist group
group_casualties <- df %>%
  group_by(Group) %>%
  summarise(Total_Casualties = sum(casualities, na.rm = TRUE)) %>%
  arrange(desc(Total_Casualties))

# Filter for the top 10 deadliest groups
top_groups <- group_casualties %>%
  top_n(10, wt = Total_Casualties)

# Plot the total casualties caused by the top 10 groups
ggplot(top_groups, aes(x = reorder(Group, -Total_Casualties), y = Total_Casualties, fill = Group)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Top 10 Deadliest Terrorist Groups by Total Casualties", x = "Terrorist Group", y = "Total Casualties") +
  theme_minimal() +
  theme(legend.position = "none")

```

below are based on smart questions
1. Which regions or countries have experienced the most attacks over the years, and how has that changed over time?

```{r}
library(dplyr)
library(ggplot2)

# Grouping the data by Year and Region
attacks_by_year_region <- df %>%
  group_by(Year, Region) %>%
  summarise(Count = n())

# Line plot of attacks over the years by region
ggplot(attacks_by_year_region, aes(x = Year, y = Count, color = Region)) +
  geom_line(size = 1.2) +
  labs(title = "Terrorist Attacks Over the Years by Region",
       x = "Year",
       y = "Count of Attacks") +
  theme_minimal() +
  scale_color_brewer(palette = "Set3") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
2. What are the most common types of attacks, and how do they vary by region or country?

```{r}
# Grouping by AttackType and Region
attacks_by_type_region <- df %>%
  group_by(AttackType, Region) %>%
  summarise(Count = n())

# Stacked bar chart
ggplot(attacks_by_type_region, aes(x = Region, y = Count, fill = AttackType)) + 
  geom_bar(stat = "identity") + 
  labs(title = "Attack Types by Region",
       x = "Region",
       y = "Count of Attacks") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set3") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```


4. Which weapon types result in the highest casualties, and are there trends in the use of different weapons over time?

```{r}
# Grouping by Year and Weapon_Type
weapon_trends <- df %>%
  group_by(Year, Weapon_Type) %>%
  summarise(Total_Casualties = sum(casualities))

# Line plot
ggplot(weapon_trends, aes(x = Year, y = Total_Casualties, color = Weapon_Type)) +
  geom_line(size = 1.2) +
  labs(title = "Casualties by Weapon Types Over Time",
       x = "Year",
       y = "Total Casualties") +
  theme_minimal() +
  scale_color_brewer(palette = "Set3") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```







